\headingfont University of Extremadura\\ Faculty of Science\\[1cm] \textbf{ Physics degree\\ Degree Final Project\\[3cm] \flushleft{\large{About implementation of LoRa-based wireless monitoring systems for tree species dendrometry.}}\\[3cm] }

### 2020-06-15 docs/paper/LaTex/Paper.tex
Completes technical description

\begin{itemize}
  \item \textbf{PoE} (Power-over-Ethernet); this is not actually a suitable option because Raspberry Pi does not support PoE by itself, it needs a \href{https://www.raspberrypi.org/products/poe-hat/}{separate HAT} which is using the GPIO connector, so this makes impossible to use it along the LoRa HAT. Apparently there are hacks to reduce this HAT size, but they also use pins in the GPIO connector \cite{RaspiPoEHack}. Besides, this method only works with up to 100m cables, so it would be necessary to have a power source and internet access point within a 100m radius.
  \item \textbf{WiFi} and battery powered; this setup implies also the existence of a relatively close power source, because the access point used to provide that wireless network connection should work also with mains power. Probably the WiFi radio is not enough to make it the difference between this setup and the next one suggested.
  \item Simple \textbf{ethernet} and mains power, this setup is similar to the previous one exposed, but omitting the WiFi limitations/configurations, the access point and the Raspberry Pi should be practically both at the same location.
  \item \textbf{GSM module} and battery or mains powered; this document consider this the ideal setup. This setup doesn't require a conventional access point because the internet access is granted through a GSM module, in a similar way than a mobile phone. This would be the ideal way because it minimizes the resources needed, which is crucial in an environment where those probably won't be available ---in the middle of the forest.

  Furthermore, this is apparently possible through two different options:
    \begin{itemize}
      \item \href{https://www.itead.cc/wiki/RPI_SIM800_GSM/GPRS_ADD-ON_V2.0}{Itead Raspberry Pi GSM Board (SIM800)}. This is not the most interesting way, because though GPIO pins comes through this does mean the HAT supports stacking, and even if that does, doesn't mean every other HAT would.
      \item \href{https://tutorials-raspberrypi.com/raspberry-pi-gsm-module-mobile-internet-lte-3g-umts/}{USB GSM module}. This is apparently the most promising option. These kind of GSM modules are in general compliant with GNU/Linux systems and using a regular SIM card, they would be eventually able to provide an internet access point. Of course, this also will to increment the size of the whole device.
    \end{itemize}
\end{itemize}

It is important to note the necessity of a relatively stable internet connection; those gateways will forward the trafic to the cloud server, so they are receiving LoRa packets through its LoRa interface and then forwarding them to the specified server.

@@

Once is available the remote control for Raspberry Pi through ssh, is possible to manage it completely from the command line, even upgrade the system and compile the required controller to get properly working the LoRa transceiver. At this point and depending on the type of access point used to provide internet service to the Raspberry, the reader will be able even to remote control the Raspberry over internet. For this purpose could be useful to read the access point documentation in order to perform port forwarding or alternatively \href{http://www.thirdway.ch/En/projects/raspberry_pi_3g/index.php}{a reverse tunnel over ssh.}\footnote{In the case the ISP does not allow ssh when using a GSM module, also the reader may find that IPv6 works but IPv4 does not. Anyway, as said the internet access point could be variable and it is not determined by this project.\\ The most important thing is to work confortably with the Raspberr Pi with no HDMI, keyboard and mouse plugged, and this can be done locally in a LAN to configure it in a frist place.}

@@

Dragino foresees this and provides a \href{https://github.com/dragino/dual_chan_pkt_fwd}{dual channel controller}. The most important thing about this dual channel controller, is not the possibility to use the transceiver in a dual channel mode, but to use it as a gateway in the physical sense.\footnote{This of course means Raspberry}

>>>>

\begin{itemize}
  \item \textbf{PoE} (Power-over-Ethernet); this is not actually a suitable option because Raspberry Pi does not support PoE by itself, it needs a \href{https://www.raspberrypi.org/products/poe-hat/}{separate HAT} which is using the GPIO connector, so this makes impossible to use it along the LoRa HAT. Apparently there are hacks to reduce this HAT size, but they also use pins in the GPIO connector \cite{RaspiPoEHack}. Besides, this method only works with up to 100m cables, so it would be necessary to have a power source and internet access point within a 100m radius.\footnote{The \href{https://github.com/hallard/LoRasPI}{LoRasPI HAT} is more suitable for this option because does not cover whole GPIO connector, leaving free the GPIO pins used in \cite{RaspiPoEHack}.}
  \item \textbf{WiFi} and battery powered; this setup implies also the existence of a relatively close power source, because the access point used to provide that wireless network connection should work also with mains power. Probably the WiFi radio is not enough to make it the difference between this setup and the next one suggested.
  \item Simple \textbf{ethernet} and mains power, this setup is similar to the previous one exposed, but omitting the WiFi limitations/configurations, the access point and the Raspberry Pi should be practically both at the same location.
  \item \textbf{GSM module} and battery or mains powered; this document consider this the ideal setup. This setup doesn't require a conventional access point because the internet access is granted through a GSM module, in a similar way than a mobile phone. This would be the ideal way because it minimizes the resources needed, which is crucial in an environment where those probably won't be available ---in the middle of the forest.

  Furthermore, this is apparently possible through two different options:
    \begin{itemize}
      \item \href{https://www.itead.cc/wiki/RPI_SIM800_GSM/GPRS_ADD-ON_V2.0}{Itead Raspberry Pi GSM Board (SIM800)}. This is not the most interesting way, because though GPIO pins comes through this does mean the HAT supports \href{http://www.pi-in-the-sky.com/index.php?id=stacking-guide}{stacking}, and even if that does, doesn't mean every other HAT would.
      \item \href{https://tutorials-raspberrypi.com/raspberry-pi-gsm-module-mobile-internet-lte-3g-umts/}{USB GSM module}. This is apparently the most promising option. These kind of GSM modules are in general compliant with GNU/Linux systems and using a regular SIM card, they would be eventually able to provide an internet access point. Of course, this also will to increment the size of the whole device.
    \end{itemize}
\end{itemize}

It is important to note the necessity of a relatively stable internet connection; those gateways will forward the traffic to the cloud server, so they are receiving LoRa packets through its LoRa interface and then forwarding them to the specified server.

@@

Once is available the remote control for Raspberry Pi through ssh, is possible to manage it completely from the command line, even upgrade the system and compile the required controller to get properly working the LoRa transceiver. At this point and depending on the type of access point used to provide internet service to the Raspberry, the reader will be able even to remote control the Raspberry over internet. For this purpose could be useful to read the access point documentation in order to perform port forwarding or alternatively \href{http://www.thirdway.ch/En/projects/raspberry_pi_3g/index.php}{a reverse tunnel over ssh.}\footnote{In the case the ISP does not allow ssh when using a GSM module, also the reader may find that IPv6 works but IPv4 does not. Anyway, as said the internet access point could be variable and it is not determined by this project.\\ The most important thing is to work comfortably with the Raspberry Pi with no HDMI, keyboard and mouse plugged, and this can be done locally in a LAN to configure it in a first place.}

@@

Dragino foresees this and provides a dual channel controller \cite{Dragino_DualChannelController_Rpi}. The most important thing about this dual channel controller, is not the possibility to use the transceiver in a dual channel mode, but to use it as a gateway in the physical sense.

In order to get the HAT operative, the reader will need to enable the SPI (Serial Peripheral Interface). This can be done also with \cmd{raspi-config}

\begin{figure}[ht]
  \centering
  \includegraphics[width=.9\textwidth]{../../pictures/SPI_raspi-config.png}
  \caption{SSH session to enable the SPI kernel module in Raspberry Pi using \cmd{raspi-config}. Inside \enquote{5 Interfacing options}}
  \label{fig:EnablingSPI}
\end{figure}

Then it would be necessary to install the GPIO access library from Raspberry repositories, \cmd{sudo apt install wiringpi}. After this, the reader will be in position to clone the controller from the github repo \cite{Dragino_DualChannelController_Rpi}

\begin{lstlisting}[%
  % float=ht,
  language=json,
  firstnumber=1,
  caption={Instructions to compile and install as a system service},
  label=ej:compiling_dual_channel_pkg
  ]
  cd ~
  git clone https://github.com/dragino/dual_chan_pkt_fwd
  cd dual_chan_pkt_fwd
  make
\end{lstlisting}

Once the controller is compiled, the reader must check the contents of \cmd{global\_conf.json}, for this projects, due it is using the LoRa GPS HAT Single Channel LoRa \cite{DraginoRpiHat}, the pins configuration in \cmd{global\_conf.json} must be

\begin{lstlisting}[%
  % float=ht,
  language=json,
  firstnumber=1,
  caption={defined pins in \cmd{global\_config.json}},
  label=ej:global_configJSON
  ]
  "pin_nss": 6,
  "pin_dio0": 7,
  "pin_rst": 0
\end{lstlisting}

I order to obtain the gateway ID the reader will need to perform the following command to run for the first time the \cmd{dual\_chan\_pkt\_fwd} binary, i.e. once the compilation process has produced the \cmd{dual\_chan\_pkt\_fwd} is necessary to run it directly to check the terminal output and find for the gateway ID.

\begin{lstlisting}[%
  % float=ht,
  language=json,
  firstnumber=1,
  caption={Running for the first time the LoRa HAT controller.},
  label=ej:Running_DualChann_Controller
  ]
  $ sudo ./dual_chan_pkt_fwd
  server: .address = router.eu.staging.thethings.network; .port = 1700; .enable = 1
  server: .address = router.eu.thethings.network; .port = 1700; .enable = 0
  Gateway Configuration
    your name (a@b.c)
    Dual channel pkt forwarder
    Latitude=0.00000000
    Longitude=0.00000000
    Altitude=10
    Interface: eth0
  Trying to detect module CE0 with NSS=6 DIO0=7 Reset=3 Led1=unused
  SX1276 detected on CE0, starting.
  Trying to detect module CE1 with NSS=6 DIO0=7 Reset=3 Led1=unused
  SX1276 detected on CE1, starting.
  Gateway ID: b8:27:eb:ff:ff:1b:14:9b
  Listening at SF7 on 868.100000 Mhz.
  Listening at SF7 on 868.100000 Mhz.
  -----------------------------------
  stat update: 2020-06-15 10:53:04 GMT no packet received yet 
\end{lstlisting}

After this last instruction (\cmd{sudo make install}) the
